Forecasting the black Sigatoka development rate: A comparison of machine learning techniques

### 2015-11-17 Paper Sigatoka - 2015.tex
lunes sigo actualizando

\title{Forecasting the black Sigatoka development rate: A machine learning methods comparison 
%\tnoteref{mytitlenote}
}
%\tnotetext[mytitlenote]{Fully documented templates are %available in the elsarticle package on \href{http://%www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}%{CTAN}.}

@@

The Black Sigatoka disease caused by the fungus {\it Mycosphaerella
  fijiensis Morelet} is the major pathological problem of banana and
plantain crops in Central America, Panama, Colombia and Ecuador, as in
many parts of Africa and Asia \citep{MarinVargas1995}.

This disease attacks the plant leaves producing a rapid deterioration
of the leaf area, affects the growth and productivity of plants as the
ability of photosynthesis decreases, causes a reduction in the quality
of the fruit, and promotes premature maturation of bunches, which is
the major cause of product losses due to this disease. \figurename
$.$\ref{figura1} shows three stages of this disease.

Phytopathological studies point out that precipitation, temperature,
relative humidity and wind are the main climatic variables that affect
the development of this disease \citep{MarinVargas1995}.

@@

According to studies by the National Banana Corporation of Costa Rica
(Corbana) made in 2013, considering on average between 53 thru 57
cycles of fungicide applications per farm, the cost per hectare per
year ranged between \$1800 USD and thru \$1900 USD. This represents
about 0.76 cents of the price of a box of 18.14 kilograms. Overall,
this represents 10\% to 12\% of the total production cost
\citet{Bresciani2015}.

The past and present disease development rate can in principle be used
to predict its future behavior, tendencies and to determine whether
particular fungicide spray schedules will be able to effectively and
economically control the disease \citet{ChuangJeger1987}.

There are efforts to apply machine learning methods for
decision-making in agriculture, including the control of crop
diseases. For example, \cite{Camargo2012} present an intelligent
systems for the assessment of crop disorders, \cite{Huang2010}
introduce a plant virus identification method based on neural networks
with an evolutionary preprocessing stage, \cite{Kim2014} summarize in
their survey crop pests prediction methods using regression and
machine learning approaches, while \cite{Zhao2013} present an
intelligent agricultural forecasting system based on wireless sensor
networks.

In this work, we compare four machine learning methods: support vector
regression (SVR), echo state networks (ESN), ridge regression, and
ordinary least squares linear regression, to predict the black
Sigatoka disease development rate.

The main contribution of this work is a comparison between machine
learning methods to forecast black Sigatoka development rate.

@@

The system measures the disease development state to determine when to
apply fungicides \citep{MarinVargas1995}.  This system is based on two
components: a climate component, which is given by the Piche
evaporation and a biological component, given by the stage of progress
or the rate of disease development. Originally, this system was
designed to work with young plants. One selected plant must exhibit a
normal growth and be in a place that enforces a healthy
development. The plant must start with 5 to 6 true leaves. The
assessments are made at fixed intervals of seven days as long as
possible, on the same plant. The first observations should consider
the leaf emission, also the level of infection on the leaves should be
evaluated considering the stages of development
\citep{MarinVargas1995}.

\subsubsection{Support Vector Regression (SVR)}
From the perspective of Support Vector Regression (SVR) the regression function $y = f(s)$ for a given dataset $D ={(s_i,y_i ) }_(i=1)^n$ , is represented as a linear function of the form (Wei, Tao, ZhuoShu, and Zio, 2013):
$f(s)=w^T s+b$
where 	$w$ and $b$ are respectively the weight vector and the intercept of the model, and they are selected to find an optimal fit to the data available in $D$.
For nonlinear cases, one proceeds by mapping the input p-dimensional vectors via a nonlinear function $ɸ∶R^p→F$, onto the feature space F.  After nonlinear mapping, the regression function evolves to a pervasive form:
$f(s)=w^T ɸ(s)+b$
SVR uses the є-insensitive loss function:
$l=|y-f(s)|_є= {█(0,   |y-f(s)|_  ≤ є@|y-f(s)|_ - є,     else)┤$
which ignores the error if the difference between the prediction value and the actual value is smaller than $є$.
$є-insensitive loss function$ allows to find the coefficients $w$ and $b$ by solving a convex optimization problem, which balances the empirical error and the generalization ability. In SVR, the empirical error is measured by the loss function є-insensitive and the generalization ability is measured by the Euclidean norm of $w$. Then, the optimization problem to identify the regression model can be formulated by (Wei, Tao, ZhuoShu, and Zio, 2013):
$minimize    J(w,ξ_i,ξ_i^* )=  1/2  ||w| |^2+ C ∑_(i=1)^n▒〖(ξ_i,ξ_i^*)〗
subject to 	{█(y_i- w^T  ɸ(s)-b  ≤   є+ξ_i@w^T  ɸ(s)+b- y_i   ≤   є+ξ_i^*     i=1,…,n@ξ_i,ξ_i^(* )≥0)┤$
where C denotes the penalty parameter between empirical and generalization errors, and  $ξ_i,ξ_i^(* )$are slack variables, as shown in Fig 2.

Fig 2: є-insensitive loss function (Wei, Tao, ZhuoShu, and Zio, 2013)
The solution of this optimization problem by the Lagrange method is:
$f(s)=w^T ɸ(s)+b= ∑_(i=1)^n▒〖(α_i-α_i^* )K(s,s_i )+b〗$
where $α_i,α_i^*$  are the Lagrange multipliers of the optimization problem’s dual form and $K(s_i,s_j )$ is the kernel function satisfying Mercer condition, and can be described by:
$K(s_i,s_j )=〈ɸ(s_i )  ɸ(s_j)〉  $
Common kernel functions are: linear, polynomial and sigmoid.
Operations in the kernel function K(s,s_i ) are performed in the input space rather than in the potentially high dimensional feature space of ɸ. An inner product in the feature space has an equivalent kernel in the input space (Alonso, Rodríguez Castañón, and Bahamonde, 2013).

\subsubsection{Ordinary least square}
This method fits a linear model with coefficients w = (w1,...,wp) to minimize the residual sum of squares between the observed responses in the dataset, and the responses predicted by the linear approximation. Mathematically it solves a problem of the form (scikit-learn developer, 2014):
〖■(min@w) 〗⁡〖‖Xw-y‖_2^2 〗

@@

This addresses some of the problems of Ordinary Least Squares by imposing a penalty on the size of coefficients. The ridge coefficients minimize a penalized residual sum of squares (scikit-learn developer, 2014):
〖■(min@w) 〗⁡〖‖Xw-y‖_2^2 〗+ α‖w‖_2^2
Here, α ˃ 0 is a complexity parameter that controls the amount of shrinkage: the larger the value of α, the greater the amount of shrinkage and thus the coefficients become more robust to collinearity.
\subsubsection{Echo State Networks (ESN)}
Recurrent Neural Networks (RNN) are useful for temporal patterns, but when they are trained with backpropagation method, they are very slow.  Echo State Network (ESN) is an alternative training method to solve that problem.  ESN is based on the observation that if a random RNN possesses certain algebraic properties, training only a linear readout from it is often sufficient to achieve excellent performance in practical applications (Lukoševičius and Jaeger, 2009). 
For a given training input signal u(n)  ϵ R^(N_u ) a desired target output signal y^(target(n))  ϵ R^(N_y )
is known. Here n = 1, . . . ,T is the discrete time and T is the number of data points in the training dataset. The task is to learn a model with output y(n)  ϵ R^(N_y ), where y(n) matches y^target(n) (n) as well as possible, minimizing an error measure E(y,y^target), and, more importantly, generalizes well to unseen data. The untrained RNN part of an ESN is called a dynamical reservoir, and the resulting states x(n) are termed echoes of its input history
 (Lukoševičius M. , 2012). Finally, these signals are sent to an output layer as shown in the following Fig.

@@

\subsubsection{Data}
In this work we used data acquired in two research farms of Corbana in Costa Rica: 1) 28 Millas (located at Matina) and La Rita (located at Pococí), both in the province of Limón, Costa Rica. The banana type is Musa AAA, subgroup Cavendish, cv. Grande Naine. Table 1 shows the variables considered initially.
Table 1 Variables used in the study
Variable	Meaning
T_(a_max )	Max air temperature
T_(a_min )	Min air temperature
T ̅_a	Mean air temperature
H	Humidity
H_min	Min humidity
H_max	Max humidity
R	Solar radiation
P ̅	Mean precipitation
W_max	Max speed wind
W ̅	Mean speed wind
L_2	Biological warning system – Leaf 2
L_3	Biological warning system – Leaf 3
L_4	Biological warning system – Leaf 4
ES	Biological warning system – Evolution Stage
The value to be predicted in all cases was ES, that is the total measure of the biological warning system.
The data on the biological warning system are collected once a week. Although Corbana has meteorological stations that take data every five minutes, for these experiments, weekly averages generated by nearby stations to each of the farms were used.
The time intervals used for this study were: La Rita, week 48 of 2002 to week 17 of the 2015 (647 weeks) and for 28 Miles, week 37 of 2003 to week 18 of 2015 (605 weeks).
Data preprocessing
In 28 Miles farm we detect that 1\% of the data were missing, while in La Rita 2.25\% of the data were missing. To complete the missing value we use spline interpolation. The data collected did not exhibit outliers.
Due the fact that the variables measure meteorological or biological process, they are discretized in order to reflect trends in the data rather than the specific continuous values. The coefficient of variation C_v (x) of each variable x was used to determine the numbers of discretizations with:
n= ⌊100 C_v (x)⌋
Each discretization range was uniformly partitioned. Besides enabling the capture of tendencies, the discretization removes the effect of small variations in the data collection, either by inaccuracies of the instruments (meteorological variables) or by subjective bias introduced by the human who collects the data (biological warning system).
Each feature was scaled in a range 0 between 1. The variable to be predicted was not scaled.

\subsubsection{Evaluation criteria}

@@

\subsubsection{Methods}
This research had two phases.
{\bf Phase one }

In the phase one, we did ten-fold-cross-validation and did a lot of proofs with different machine learning methods and different configuration. 
We proved with several combinations:
	Patterns: From one week of observed data to predict the next week until nine weeks before to predict two weeks later.
	Algorithms: Support vector regression with different kernel functions: linear, RBF (Gaussian) and sigmoid; echo state networks; ordinary least squares linear regression and ridge regression.
	Variables included in the model. We proved the following combinations:
	All variables.
	Only variables that according to expert judgment have more impact on the black Sigatoka development: humidity, precipitation, temperature and wind speed (Marin Vargas and Romero Calderón, 1995).
	From the four variables listed in the previous paragraph, runs were conducted using each of the variables separately, and combining other runs all the possible pairs of those four variables.
Phase two
In the second phase, we used the best configurations obtained en la phase one and did validation with the last 52 and 102 weeks. 
This second phase pretended to show how these methods behaved on a time of important climate change how are 2014 and 2015 years.
Programming environment
We use python programming language with the Integrated Development Environment (IDE) Spyder, particularly with libraries: pandas (Comunity, 2014); numpy (numpy.org, 2013); for SVR, ridge and ordinary least squares, we used sklearn (Pedregosa, et al., 2011); and for ESN the python-based code used belongs to Dr. Mantas Lukoševičius (2012) from which we made the necessary adjustments for the experiments of this research. The computer was a Lenovo ThinkPad, processor Intel(R) Core™ i7-4800MQ CPU @ 2.70GHz, 16.0 GB RAM, running Windows 8 Pro.

>>>>

\title{Forecasting the black Sigatoka development rate: A comparison of machine learning techniques 
%\tnoteref{mytitlenote}
}
%\tnotetext[mytitlenote]{Fully documented templates are %available in the elsarticle package on \href{http://%www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}%{CTAN}.}

@@

The black Sigatoka disease caused by the fungus {\it Mycosphaerella
  fijiensis Morelet} is the major pathological problem of banana and
plantain crops in Central America, Panama, Colombia and Ecuador, as in
many parts of Africa and Asia \citep{MarinVargas1995}.

This disease attacks the plant leaves producing a rapid deterioration
of the leaf area, affects the growth and productivity of the plants due to the impairment of their photosinthetic ability  causes a reduction in the quality
of the fruit, and promotes premature maturation of bunches, which is
the major cause of product losses associated with the black Sigatoka. \figurename
$.$\ref{figura1} shows three stages of this disease.

Phytopathological studies point out that precipitation, temperature,
relative humidity and wind are the main climatic variables that affect
its development \citep{MarinVargas1995}.

@@

According to studies by the National Banana Corporation of Costa Rica
(Corbana) made in 2013, considering on average between 53 thru 57
cycles of fungicide applications per farm, the cost per hectare per
year ranged between \$1800 USD and \$1900 USD. This represents
about 0.76 cents of the price of a box of 18.14 kilograms. Overall,
this represents 10\% to 12\% of the total production cost
\citet{Bresciani2015}.

The past and present rates of disease development  can in principle be used
to predict its future behavior and to determine whether
particular fungicide spray schedules will be able to effectively and
economically control the disease \citet{ChuangJeger1987}.

There are efforts to apply machine learning methods for
decision-making in agriculture, including the control of crop
diseases. For example, \cite{Camargo2012} present an intelligent
system for the assessment of crop disorders, \cite{Huang2010}
introduce a plant virus identification method based on neural networks
with an evolutionary preprocessing stage, \cite{Kim2014} summarize in
their survey crop pests prediction methods using regression and
machine learning approaches, while \cite{Zhao2013} present an
intelligent agricultural forecasting system based on wireless sensor
networks.

In this work, we compare four machine learning techniques (support vector
regression (SVR), echo state networks (ESN), ridge regression and
ordinary least squares linear regression) to predict the development rate of the black Sigatoka disease.

The main contribution of this work is a comparison between machine
learning methods to forecast black Sigatoka development rate. (FALTA COMPLETAR)

@@

This system measures the disease development state to determine when to
apply fungicides \citep{MarinVargas1995}.  This system is based on two
components: a climate component, which is given by the Piche
evaporation and a biological component, given by the stage of progress
or the rate of disease development. Originally, this system was
designed to work with young plants. One selected plant must exhibit a
normal growth and be in a place that enforces a healthy
development. The plant must start with 5 to 6 true leaves. The
assessments are made at fixed intervals of seven days as long as
possible, on the same plant. The first observations should consider
the leaf emission, also the level of infection on the leaves should be
evaluated considering the stages of development
\citep{MarinVargas1995}.

\subsubsection{Support Vector Regression (SVR)}
From the perspective of Support Vector Regression (SVR) the regression function $y = f(s)$ for a given dataset $D ={(s_i,y_i ) }_(i=1)^n$ , is represented as a linear function of the form (Wei, Tao, ZhuoShu, and Zio, 2013):
$f(s)=w^T s+b$
where 	$w$ and $b$ are respectively the weight vector and the intercept of the model, and they are selected to find an optimal fit to the data available in $D$.
For nonlinear cases, one proceeds by mapping the input p-dimensional vectors via a nonlinear function $ɸ∶R^p→F$, onto the feature space F.  After nonlinear mapping, the regression function evolves to a pervasive form:
$$f(s)=w^T ɸ(s)+b$$
SVR uses the є-insensitive loss function:
% $$l=|y-f(s)|_є= {(0,   |y-f(s)|_  ≤ є |y-f(s)|_ - є,     else)$$

which ignores the error if the difference between the prediction value and the actual value is smaller than $є$.
$є-insensitive loss function$ allows to find the coefficients $w$ and $b$ by solving a convex optimization problem, which balances the empirical error and the generalization ability. In SVR, the empirical error is measured by the loss function є-insensitive and the generalization ability is measured by the Euclidean norm of $w$. Then, the optimization problem to identify the regression model can be formulated by (Wei, Tao, ZhuoShu, and Zio, 2013):
% $minimize    J(w,ξ_i,ξ_i^* )=  1/2  ||w| |^2+ C ∑_(i=1)^n▒〖(ξ_i,ξ_i^*)〗
% subject to 	{█(y_i- w^T  ɸ(s)-b  ≤   є+ξ_i@w^T  ɸ(s)+b- y_i   ≤   є% 
% +ξ_i^*     i=1,…,n@ξ_i,ξ_i^(* )≥0)┤$
where C denotes the penalty parameter between empirical and generalization errors, and  $ξ_i,ξ_i^(* )$are slack variables, as shown in Fig 2.

Fig 2: є-insensitive loss function (Wei, Tao, ZhuoShu, and Zio, 2013)
The solution of this optimization problem by the Lagrange method is:
% $f(s)=w^T ɸ(s)+b= ∑_(i=1)^n▒〖(α_i-α_i^* )K(s,s_i )+b〗$
where $α_i,α_i^*$  are the Lagrange multipliers of the optimization problem’s dual form and $K(s_i,s_j )$ is the kernel function satisfying Mercer condition, and can be described by:
% $K(s_i,s_j )=〈ɸ(s_i )  ɸ(s_j)〉  $
Common kernel functions are: linear, polynomial and sigmoid.
Operations in the kernel function $K(s,s_i $) are performed in the input space rather than in the potentially high dimensional feature space of $ɸ$. An inner product in the feature space has an equivalent kernel in the input space (Alonso, Rodríguez Castañón, and Bahamonde, 2013).

\subsubsection{Ordinary least square}
This method fits a linear model with coefficients w = (w1,...,wp) to minimize the residual sum of squares between the observed responses in the dataset, and the responses predicted by the linear approximation. Mathematically it solves a problem of the form (scikit-learn developer, 2014):
% 〖■(min@w) 〗⁡〖‖Xw-y‖_2^2 〗

@@

This addresses some of the problems of Ordinary Least Squares by imposing a penalty on the size of coefficients. The ridge coefficients minimize a penalized residual sum of squares (scikit-learn developer, 2014):
% 〖■(min@w) 〗⁡〖‖Xw-y‖_2^2 〗+ α‖w‖_2^2
Here, α ˃ 0 is a complexity parameter that controls the amount of shrinkage: the larger the value of α, the greater the amount of shrinkage and thus the coefficients become more robust to collinearity.
\subsubsection{Echo State Networks (ESN)}
Recurrent Neural Networks (RNN) are useful for temporal patterns, but when they are trained with backpropagation method, they are very slow.  Echo State Network (ESN) is an alternative training method to solve that problem.  ESN is based on the observation that if a random RNN possesses certain algebraic properties, training only a linear readout from it is often sufficient to achieve excellent performance in practical applications (Lukoševičius and Jaeger, 2009). 
For a given training input signal $u(n)  ϵ R^(N_u )$ a desired target output signal 
% $y^(target(n))  ϵ R^(N_y )$
is known. Here $n = 1, . . . ,T$ is the discrete time and $T$ is the number of data points in the training dataset. The task is to learn a model with output $y(n)  ϵ R^(N_y )$, where $y(n)$ matches $y^target(n)$ $(n)$ as well as possible, minimizing an error measure $E(y,y^target)$, and, more importantly, generalizes well to unseen data. The untrained RNN part of an ESN is called a dynamical reservoir, and the resulting states x(n) are termed echoes of its input history
 (Lukoševičius M. , 2012). Finally, these signals are sent to an output layer as shown in the following Fig.

@@

\subsection{Data}
In this work we used data acquired in two research farms of Corbana in Costa Rica: 1) 28 Millas (located at Matina) and La Rita (located at Pococí), both in the province of Limón, Costa Rica. The banana type is Musa AAA, subgroup Cavendish, cv. Grande Naine. Table 1 shows the variables considered initially.
Table 1 Variables used in the study
Variable	Meaning
%
%T_(a_max )	Max air temperature
%T_(a_min )	Min air temperature
%T ̅_a	Mean air temperature
%H	Humidity
%H_min	Min humidity
%H_max	Max humidity
%R	Solar radiation
%P ̅	Mean precipitation
%W_max	Max speed wind
%W ̅	Mean speed wind
%L_2	Biological warning system – Leaf 2
%L_3	Biological warning system – Leaf 3
%L_4	Biological warning system – Leaf 4
%ES	Biological warning system – Evolution Stage

The value to be predicted in all cases was ES, that is the total measure of the biological warning system.
The data on the biological warning system are collected once a week. Although Corbana has meteorological stations that take data every five minutes, for these experiments, weekly averages generated by nearby stations to each of the farms were used.
The time intervals used for this study were: La Rita, week 48 of 2002 to week 17 of the 2015 (647 weeks) and for 28 Miles, week 37 of 2003 to week 18 of 2015 (605 weeks).
Data preprocessing
In 28 Miles farm we detect that 1\% of the data were missing, while in La Rita 2.25\% of the data were missing. To complete the missing value we use spline interpolation. The data collected did not exhibit outliers.
Due the fact that the variables measure meteorological or biological process, they are discretized in order to reflect trends in the data rather than the specific continuous values. The coefficient of variation $C_v (x)$ of each variable x was used to determine the numbers of discretizations with:
$$n= ⌊100 C_v (x)⌋$$
Each discretization range was uniformly partitioned. Besides enabling the capture of tendencies, the discretization removes the effect of small variations in the data collection, either by inaccuracies of the instruments (meteorological variables) or by subjective bias introduced by the human who collects the data (biological warning system).
Each feature was scaled in a range 0 between 1. The variable to be predicted was not scaled.

\subsection{Evaluation criteria}

@@

\subsection{Methods}
This research had two phases.\\
{\bf Phase one } \\

In the phase one, we did ten-fold-cross-validation and did a lot of proofs with different machine learning methods and different configuration. 
We proved with several combinations:
	Patterns: From one week of observed data to predict the next week until nine weeks before to predict two weeks later.
	Algorithms: Support vector regression with different kernel functions: linear, RBF (Gaussian) and sigmoid; echo state networks; ordinary least squares linear regression and ridge regression.
	Variables included in the model. We proved the following combinations:
	All variables.
	Only variables that according to expert judgment have more impact on the black Sigatoka development: humidity, precipitation, temperature and wind speed (Marin Vargas and Romero Calderón, 1995).
	From the four variables listed in the previous paragraph, runs were conducted using each of the variables separately, and combining other runs all the possible pairs of those four variables.\\
{\bf Phase two } \\
In the second phase, we used the best configurations obtained en la phase one and did validation with the last 52 and 102 weeks. 
This second phase pretended to show how these methods behaved on a time of important climate change how are 2014 and 2015 years.

Programming environment
We use python programming language with the Integrated Development Environment (IDE) Spyder, particularly with libraries: pandas (Comunity, 2014); numpy (numpy.org, 2013); for SVR, ridge and ordinary least squares, we used sklearn (Pedregosa, et al., 2011); and for ESN the python-based code used belongs to Dr. Mantas Lukoševičius (2012) from which we made the necessary adjustments for the experiments of this research. The computer was a Lenovo ThinkPad, processor Intel(R) Core™ i7-4800MQ CPU @ 2.70GHz, 16.0 GB RAM, running Windows 8 Pro.
